\documentclass[12pt,twoside,cover,color,table]%
  {fithesis3/fithesis3/fithesis3} % 0_0

% Packages
\usepackage[english]{babel} % Babel
\usepackage[utf8]{inputenc} % UTF-8
\usepackage{hologo}         % \XeLaTeX{} and friends
\usepackage[                % Microtype
  protrusion                  % Protrusion support
]{microtype}
\usepackage{amsmath}        % The split and multline environments
\usepackage{setspace}       % Paragraph spacing
\usepackage{blindtext}      % Blind text
\usepackage[toc,page]%      % Appendices
  {appendix}
\usepackage{csquotes}       % Quotations
\usepackage[                % Colors
  usenames,dvipsnames,
  svgnames,table
]{xcolor}
\usepackage[                % Fonts
  scaled=1.03
]{inconsolata}

\usepackage[                % Hyperref
  pdfborderstyle={/S/U/W 1},  % Less obtrusive borders
  hyperfootnotes,             % Footnote hyperlinks
  breaklinks                  % Better cross-page link handling
]{hyperref}

\usepackage{pgfplots}       % Graphs
\pgfplotsset{compat=1.8}
\usepgfplotslibrary{statistics}

% Color definitions
\definecolor{pantone122}{HTML}{FFD451}
\definecolor{pantone300}{HTML}{0067C6}
\definecolor{cured}{RGB}{210,45,64}
\usepackage[
  labelfont=bf              % Typeset table captions in bold font
]{caption}
\usepackage{multirow}       % Multirows
\usepackage{./definitions}  % My own macros

% Fithesis options
\thesissetup{
  title    = The Form of Theses Written in LaTeX,
  TeXtitle = The Form of Theses\\ Written in \LaTeX{},
  author   = \VN,
  advisor  = {doc. RNDr. \PS, Ph.D.},
  basepath = ./fithesis3/fithesis3,
  assignment = {./misc/declaration.pdf, 1, {},
                ./misc/assignment.pdf,  1, {}}}

\thesislong{thanks}{I wish to express my most sicere gratitude and
  appreciation to \thesis{advisor} for sharing expertise, valuable
  guidance and encouragement, doc. Ing. Michal Brandejs, CSc. for
  providing me with statistical data, which proved invaluable for
  the initial research, and one and all who, directly or
  indirectly, have lent their hand in this venture. I also thank my
  parents for their unceasing encouragement and support during my
  studies.}

% An accessor to the \thesis@ macro
\makeatletter
\def\thesis#1{\makeatletter\thesis@{#1}\makeatother}
\makeatother

% Bibliography
\usepackage[
  backend=biber,
  style=numeric,
  citestyle=numeric-comp,
  sorting=none
]{biblatex}

\addbibresource{database.bib}
%\renewcommand*{\bibfont}{\footnotesize}

% Index
\usepackage{makeidx}
\makeindex

% Glossary and acronyms
\usepackage[xindy,acronym,toc]{glossaries}
\makeglossaries
\input{glossaries}
\setglossarystyle{altlistgroup}

% Penalties
%% Zero tolerance for widows and orphans
\clubpenalty=10000

% Miscellaneous
%% Enforce more dense float-pages
\renewcommand{\floatpagefraction}{0.7}

\begin{document}

\chapter{Introduction}
\chapter{Existing Codebase}\label{sec:init-state}
  The implementation part of this thesis concerns the development
  of a new version of an already existing software project. In this
  chapter, the author is going to briefly describe the history of
  the project prior to his involvement, the state of the project's
  codebase at the time of his arrival as well as his contributions,
  which aren't directly related to the \textsf{fithesis3}
  \gls{docclass}, but bear relevance to its development.

  \section{The \textsf{fithesis1} and \textsf{fithesis2}
    classes}
  The \textsf{fithesis1} \gls{docclass} had been created by Mgr.
  Daniel Marek in 1998 as a bachelor's project under the
  supervision of \thesis{advisor}. Despite the best effort on part
  of both the author and the staff of the Computer Systems Unit at
  \gls{fi}, neither the source code of \textsf{fithesis1}, nor the
  thesis accompanying the project, was recovered.

  In 2002, the main maintainership was transferred to RNDr. Jan
  Pavlovič, Ph.D. who extended the class to serve as one of the
  backends to his \gls{docbook}-based system for thesis writing.
  This project is described separately in Section \ref{sec:xslt}.
  During 2002--2008, the codebase of \textsf{fithesis1} remained
  relatively stable under the maintenance of Jan Pavlovič and Petr
  Sojka, with one of the few major additions being the addition of
  support for Slovak by Mgr. Peter Čerenský in 2004.

  In 2008, Mgr. Stanislav Filipčík created a derivative of
  \textsf{fithesis1} called \textsf{fithesis2} as the
  implementation part of his bachelor's thesis \cite{Filipcik09}.
  This derivative contained desirable changes, such as the removal
  of the mandatory ISO Latin-2 \gls{charenc} and the addition of
  the logos of all the faculties of \gls{mu}. These changes,
  however, were insufficiently documented, which is why
  \textsf{fithesis2} wasn't released. In 2012, Bc. Tomáš Janoušek
  was tasked to document the code of \textsf{fithesis2}, which he
  did only partially \cite{fithesis2Code}.

  During January--March 2015, the author and Petr Sojka finished
  the technical documentation of \textsf{fithesis2}, refactored its
  source code and published it on the website of \gls{fi}
  \cite{fithesisWeb}. A thematic IS forum \cite{fithesisForum} was
  also founded by doc. Ing. Michal Brandejs, CSc. to foster
  discussion about the typesetting of theses at the Masaryk
  University.

  \section{The \textsf{xslt} and \textsf{xslt2} modules}
  \label{sec:xslt}
  The \textsf{xslt} module was developed in 2001 by RNDr. Jan
  Pavlovič, Ph.D. as the implementation part of his bachelor's
  thesis \cite{PavlovicO1}. The module combined several tools to
  enable the conversion of theses written using the \gls{docbook}
  format into various output formats including the
  \textsf{fithesis1} class.  The development of the module
  continued until 2003. During 2003--2004, Jan Pavlovič co-wrote
  several articles concerning \gls{docbook} \cite{Pavlovic03,
  PitPav03, PitPav04} and released a rewritten version of the
  module called \textsf{xslt2}. The latest contribution to the
  \textsf{xslt2} module was the implementation part of the
  bachelor's thesis of Bc. Tomáš Baluch \cite{Baluch09} supervised
  by Jan Pavlovič. The thesis modernized the user interface of the
  module and added support for access to Java content repositories.

  During January--March 2015, the author tested the final version
  of \textsf{fithesis1} with the final version of \textsf{xslt2}
  available on the UNIX machines at \gls{fi} and concluded that the
  projects retained compatibility and could be used without
  apparent issues. The author therefore included a section
  regarding the use of the \textsf{xslt2} module into the
  documentation of \textsf{fithesis1} and restored the website
  dedicated to the module \cite{xslt2web}. Since the \textsf{xslt2}
  module is only loosely coupled with the \textsf{fithesis1}
  \gls{docclass}, the author also concluded that the module should
  be relatively simple to extend to support \textsf{fithesis2} and
  \textsf{fithesis3}.

\chapter{Research}
  During the course of December 2014 and January 2015, the author
  performed an initial research to obtain user feedback to the
  \textsf{fithesis1} and \textsf{fithesis2} \glspl{docclass}, to
  analyze the trends in the use of \gls{tex} across the faculties
  of \gls{mu} and to assess the existing solutions for the
  typesetting of theses. In this chapter, the author is going to
  summarize the results and findings of his research.

  \section{User Survey}
  At the end of December 2014, an online questionnaire was
  distributed amongst the students of \gls{mu}. Table
  \ref{table:survey-faculty} illustrates the distribution of the
  respondents across the faculties of \gls{mu} according to the
  claims of the respondents. It should be noted that several
  respondents claimed to be studying at more than one faculty of
  \gls{mu}, which is the reason for the distortion of the
  percentage. The raw data obtained from the questionnaire in Czech
  is available in the \texttt{survey/survey.csv} file distributed
  alongside this thesis.

  \begin{table}[!bt]
    \caption{The distribution of the questionnaire
      respondents across the faculties of \acrshort{mu}}
    \begin{tabularx}{\textwidth}{Xcr} \textbf{On which faculty of
      \gls{mu} do you study?} & \textbf{\#} & \textbf{\%}
      \\ \toprule Faculty of Informatics                  & $82$
      & $92.1$ \\ Faculty of Science                      &  $3$
      &  $3.4$ \\ Faculty of Education                    &  $2$
      &  $2.2$ \\ Faculty of Social Studies               &  $2$
      &  $2.2$ \\ Faculty of Law                          &  $1$
      &  $1.1$ \\ Faculty of Medicine                     &  $1$
      &  $1.1$ \\ Faculty of Arts                         &  $1$
      &  $1.1$ \\ Faculty of Economics \& Administration  &  $1$
      &  $1.1$ \\ Faculty of Sports Studies               &  $0$
      &  $0.0$ \\ \bottomrule \textbf{Total}      &$\mathbf{89}$
      & $\mathbf{100.0}$
    \end{tabularx}
    \label{table:survey-faculty}
  \end{table}

  The overwhelming majority of respondents claimed that the highest
  degree for which they were studying was that of a bachelor (see
  Table \ref{table:survey-type}) and that they were planning to use
  \gls{tex} to write their theses (see Table
  \ref{table:survey-sw}). Most of those who claimed to be planning
  to use \gls{tex} also claimed to know about the existence of the
  \textsf{fithesis1} \gls{docclass} family and claimed to be
  planning to use it as a template for their theses (see Table
  \ref{table:survey-tex}).

  \begin{table}
    \caption{The highest academic degrees currently pursued by the
      respondents of the questionnaire}
    \begin{tabularx}{\textwidth}{Xcr}
      \textbf{Which academic degree are you currently pursuing?} &
      \textbf{\#} & \textbf{\%} \\
      \toprule
      Bachelor's degree  & $70$          & $78.7$ \\
      Master's degree    & $17$          & $19.1$ \\
      Doctorate          & $2$           & $2.2$  \\
      \bottomrule
      \textbf{Total}     & $\mathbf{89}$ & $\mathbf{100.0}$
    \end{tabularx}
    \label{table:survey-type}
  \end{table}

  \begin{table}
    \caption{The software the respondents of the questionnaire
      were using or planning to use to write their theses}
    \begin{tabularx}{\textwidth}{Xcr}
      \textbf{Which application do you use / are you planning to
      use to write your thesis?} & \textbf{\#} & \textbf{\%} \\
      \toprule
      \TeX{ / \LaTeX{}}     & $65$          & $73.0$ \\
      Microsoft Office      & $16$          & $18.0$ \\
      Apache OpenOffice, LibreOffice or another free office
      software suite        & $4$           &  $4.5$ \\
      Other                 & $4$           &  $4.5$ \\
      Google Documents      & $0$           &  $0.0$ \\
      \bottomrule
      \textbf{Total}        & $\mathbf{89}$ & $\mathbf{100.0}$
    \end{tabularx}
    \label{table:survey-sw}
  \end{table}

  \begin{table}
    \caption{The attitude towards using the \textsf{fithesis1}
      \gls{docclass} amongst those respondents who claimed to
      be using or planning to use \gls{tex} to typeset their theses}
    \begin{tabularx}{\textwidth}{Xcr}
      \textbf{Are you planning to use the \gls{mu}
        \textsf{fithesis1} \LaTeX{} class?} & \textbf{\#} &
        \textbf{\%} \\
      \toprule    Yes                                        & $47$
      & $72.3$ \\ Maybe, I did not know it existed           & $10$
      & $15.4$ \\ No, I'm going to use another \gls{docclass} & $5$
      &  $7.7$ \\ Other                                       & $2$
      &  $3.1$ \\ No, I'm going to use \gls{plain}            & $1$
      &  $1.5$ \\ \bottomrule
      \textbf{Total}             & $\mathbf{65}$ & $\mathbf{100.0}$
    \end{tabularx}
    \label{table:survey-tex}
  \end{table}

  Some of the respondents who claimed to be using or planning to
  use \gls{tex} to typeset their theses also provided feedback
  regarding the \textsf{fithesis1} class. Unknown to the author at
  the time of the survey was the fact that \textsf{fithesis2} had
  never been made publicly available as detailed in Section
  \ref{sec:init-state}. As a result of that, much of the feedback
  was regarding issues long fixed in \textsf{fithesis2}. Among
  the feedback relevant to \textsf{fithesis2} were calls for a more
  extensive user and technical documentation\implemented{} and a
  suggestion that the class should support the typesetting of
  printed and electronic versions of theses as separate
  documents\partimp{}.

  \section{Statistical analysis of existing theses}
  Along with the user survey, a statistical analysis of theses
  defended at \gls{mu} during 2010--2015 was carried out by the
  author. The sample data for the analysis were kindly provided by
  doc. Ing. Michal Brandejs, CSc., the head of the Computer Systems
  Unit at \gls{fi}. The raw data is available within the directory
  \texttt{statistics/} distributed alongside this thesis.

  \begin{table}
    \caption{The distribution of theses defended between 2010 and
      2015 across the faculties of \acrshort{mu}}
    \begin{tabularx}{\textwidth}{Xrr}
      \textbf{Faculty} & \textbf{\#} & \textbf{\%} \\
      \toprule
      Arts                         & $10\,000$ & $21.98$ \\% 1421
      Education                    & $8\,219$  & $18.07$ \\% 1441
      Social Studies               & $5\,599$  & $12.31$ \\% 1423
      Science                      & $5\,275$  & $11.60$ \\% 1431
      Law                          & $4\,824$  & $10.60$ \\% 1422
      Economics \& Administration  & $4\,591$  & $10.09$ \\% 1456
      Informatics                  & $2\,904$  &  $6.38$ \\% 1433
      Medicine                     & $2\,014$  &  $4.43$ \\% 1411
      Sports Studies               & $2\,062$  &  $4.53$ \\% 1451
      \bottomrule
      \textbf{Total}     & $\mathbf{45\,488}$ & $\mathbf{100.00}$
    \end{tabularx}
    \label{table:statistics-faculty}
  \end{table}

%  \begin{table}
%    \caption{The distribution of theses defended between 2010 and
%      2015 across the study programme degrees}
%    \begin{tabularx}{\textwidth}{Xrr}
%      \textbf{Degree} & \textbf{\#} & \textbf{\%} \\
%      \toprule
%      Bachelor's & $22\,288$ & $49.00$ \\
%      Master's   & $20\,761$ & $45.64$ \\
%      Doctoral   &  $2\,231$ &  $4.90$ \\
%      Lifelong   &     $208$ &  $0.46$ \\
%      \bottomrule
%      \textbf{Total} & $\mathbf{45\,488}$ & $\mathbf{100.00}$
%    \end{tabularx}
%    \label{table:statistics-degree}
%  \end{table}

%  Tables \ref{table:statistics-faculty} and
%  \ref{table:statistics-degree} detail the distribution of theses
%  written and defended during 2010--2015 across the faculties
%  of \gls{mu} and across the study programme degrees, respectively.
  Table \ref{table:statistics-faculty} detail the distribution of
  theses written and defended during 2010--2015 across the
  faculties of \gls{mu} and Table \ref{table:statistics-tex}
  illustrates how many of these theses were written using
  \gls{tex}. Table \ref{table:statistics-tex-yearly} then details
  the trends in the usage of \gls{tex} by the students of
  bachelor's, master's and doctoral degree programmes at \gls{fi}
  and \gls{sci}. Other faculties of \gls{mu} were not considered,
  since the number of theses written at them using \gls{tex} was
  statistically insignificant (see Table
  \ref{table:statistics-tex}). Theses written by students of
  lifelong education programmes were likewise ignored, since none
  of them were written using \gls{tex}.

  \begin{table}
    \caption{The distribution of theses written using \gls{tex},
      which were defended during 2010--2015 across the faculties of
      \acrshort{mu}}
    \begin{tabularx}{\textwidth}{Xrrr}
      \textbf{Faculty} & \textbf{With \gls{tex}} & \textbf{Total} &
      \textbf{\%} \\
      \toprule
      Informatics                 & $1\,716$ & $2\,904$  &
      $59.09$ \\% 1433
      Science                     & $786$     & $5\,275$  &
      $14.90$ \\% 1431
      Economics \& Administration & $64$      & $4\,591$  &
      $1.39$  \\% 1456
      Arts                        & $69$      & $10\,000$ &
      $0.69$  \\% 1421
      Medicine                    & $8$       & $2\,014$  &
      $0.40$  \\% 1411
      Law                         & $15$      & $4\,824$  &
      $0.31$  \\% 1422
      Education                   & $19$      & $8\,219$  &
      $0.23$  \\% 1441
      Social Studies              & $12$      & $5\,599$  &
      $0.21$  \\% 1423
      Sports Studies              & $3$       & $2\,062$  &
      $0.15$  \\% 1451
      \bottomrule
      \textbf{Total} & $\mathbf{2\,692}$ & $\mathbf{45\,488}$ &
      $\mathbf{5.92}$
    \end{tabularx}
    \label{table:statistics-tex}
  \end{table}

  \begin{table}
    \caption{The percentage of theses written using \gls{tex} which
      were defended in each year during 2010--2014 and the
      sample correlation coefficient $R$ between the percentage and
      the years. Remarkably strong correlations are emphasized.}
    \begin{tabularx}{\textwidth}{Xlrrrrrr}
      \textbf{Degree} & \textbf{Fac.} & \textbf{2010} &
      \textbf{2011} & \textbf{2012} & \textbf{2013} & \textbf{2014}
      & $R$\\
      \toprule
      Bachelor's
        & \acrshort{fi}  & $58.92$ & $59.44$ & $49.54$ & $53.77$ &
          $59.06$ & \textcolor{red}{$-0.195$} \\
        & \acrshort{sci} & $11.55$ & $13.00$ & $15.90$ & $19.79$ &
          $15.16$ & \textcolor{OliveGreen}{$+0.703$} \\
        & All & $5.08$ & $6.19$ &
          $6.00$ & $6.08$ & $6.24$ &
          \textcolor{OliveGreen}{$+0.731$} \\ \midrule
      Master's
        & \acrshort{fi}  & $60.61$ & $59.91$ & $60.08$ & $64.50$ &
          $57.96$ & \textcolor{red}{$-0.046$} \\
        & \acrshort{sci} & $19.38$ & $13.54$ & $13.75$ & $13.78$ &
          $17.71$ & \textcolor{red}{$-0.180$} \\
        & All & $6.02$ & $4.88$ &
          $5.22$ & $6.59$ & $6.29$ &
          \textcolor{OliveGreen}{$+0.490$} \\ \midrule
      Doctoral
        & \acrshort{fi}  & $100.00$ & $76.67$ & $71.88$ & $83.87$ &
          $90.91$ & \textcolor{red}{$-0.155$} \\
        & \cellemph\acrshort{sci} & \cellemph$18.09$ &
          \cellemph$10.71$ & \cellemph$12.75$ & \cellemph$10.19$ &
          \cellemph$8.85$ & \cellemph\textcolor{red}{$-0.830$} \\
        & All & $8.83$ & $8.23$ &
          $8.41$ & $9.38$ & $7.43$ &
          \textcolor{red}{$-0.361$} \\
        \bottomrule
      \textbf{All}
        & \textbf{\acrshort{fi} } & $\mathbf{60.83}$ &
          $\mathbf{60.53}$ & $\mathbf{54.92}$ & $\mathbf{60.57}$ &
          $\mathbf{59.34}$ & \textcolor{red}{$\mathbf{-0.188}$} \\
        & \textbf{\acrshort{sci}} & $\mathbf{14.86}$ &
          $\mathbf{12.96}$ & $\mathbf{14.74}$ & $\mathbf{16.55}$ &
          $\mathbf{15.45}$ &
          \textcolor{OliveGreen}{$\mathbf{+0.577}$}\\
        & \cellemph\textbf{All} & \cellemph$\mathbf{5.67}$ &
          \cellemph$\mathbf{5.70}$ & \cellemph$\mathbf{5.73}$ &
          \cellemph$\mathbf{6.41}$ & \cellemph$\mathbf{6.28}$ &
          \cellemph\textcolor{OliveGreen}{$\mathbf{+0.855}$}
    \end{tabularx}
    \label{table:statistics-tex-yearly}
  \end{table}

  A thesis was considered to be written using \gls{tex}, if one
  or more files submitted with it satisfied one or more of the
  following conditions: \begin{itemize}
    \item The suffix was \texttt{tex}.
    \item The \gls{magnum} was that of a \acrshort{dvi} file.
    \item The \gls{mime} was \texttt{application/postscript} and
      the file contained the \texttt{TeXDict} substring suggesting
      that the file was a \gls{ps} document, which had been created
      using the \textsf{dvips} utility.
    \item The \gls{mime} was \texttt{application/pdf} and either
      the \texttt{Creator} or the \texttt{Producer} \gls{pdf}
      header contained the \texttt{TeX} substring suggesting that
      the file had been created using either the \textsf{dvipdfm}
      utility or a \gls{texeng}, which supports \gls{pdf} output.
  \end{itemize} Provided the heuristic is sound, there has
  been a marked and
  steady increase in the use of \gls{tex} for the typesetting of
  theses during 2010--2014 (see Table
  \ref{table:statistics-tex-yearly}). This, however, doesn't
  necessarily hold true for individual faculties and degree study
  programmes with some of them showing barely any correlation
  between the years and the use of \gls{tex} and others showing a
  strong negative correlation. A particularly striking example of
  the latter is the pronounced downwards trend in the use of
  \gls{tex} for the typesetting of doctoral theses at \gls{sci}.

%  \begin{table}
%    \caption{The sample means $M_1,M_2$ and the sample standard
%      deviations $S_1,S_2$ of the grades given to theses written
%      and defended during 2010--2015 at \gls{fi}, \gls{sci} and all
%      the faculties of \gls{mu} using and not using \gls{tex},
%      respectively}
%    \begin{tabularx}{\textwidth}{XXrrrr}
%        \bf Faculty & \bf Degree & $\mathbf{M_1}$ & $\mathbf{M_2}$
%        & $\mathbf{S_1}$ & $\mathbf{S_2}$ \\
%      \toprule
%      Informatics  % 1433
%        & Bachelor's & $2.360$ & $3.210$ & $1.498$ & $1.691$ \\
%        & Master's   & $2.379$ & $3.004$ & $1.642$ & $1.764$ \\
%        & Doctoral   & $1.639$ & $1.708$ & $1.370$ & $1.654$ \\
%        & \textbf{All} & $\mathbf{2.323}$ & $\mathbf{3.103}$
%          \bf & $\mathbf{1.566}$ & $\mathbf{1.740}$ \\
%
%      Science % 1431
%        & Bachelor's & $1.978$ & $2.066$ & $1.226$ & $1.226$ \\
%        & Master's   & $1.903$ & $2.041$ & $1.333$ & $1.300$ \\
%        & Doctoral   & $1.406$ & $1.633$ & $0.849$ & $1.154$ \\
%        & \textbf{All} & $\mathbf{1.902}$ & $\mathbf{2.010}$
%          & $\mathbf{1.251}$ & $\mathbf{1.254}$ \\ \bottomrule
%
%      \textbf{All faculties}
%      & \textbf{Bachelor's} & $\mathbf{2.235}$ & $\mathbf{2.530}$ &
%        $\mathbf{1.422}$ & $\mathbf{1.479}$ \\
%      & \textbf{Master's} & $\mathbf{2.234}$ & $\mathbf{2.312}$ &
%        $\mathbf{1.552}$ & $\mathbf{1.447}$ \\
%      & \textbf{Doctoral} & $\mathbf{1.576}$ & $\mathbf{1.849}$ &
%        $\mathbf{1.250}$ & $\mathbf{1.420}$ \\
%      & \textbf{All}    & $\mathbf{2.118}$ & $\mathbf{2.396}$ &
%        $\mathbf{1.478}$ & $\mathbf{1.471}$ \\
%    \end{tabularx}
%    \label{table:statistics-tex-grades}
%  \end{table}

  \begin{table}
    \caption{The contingency table of the numbers of marks awarded
      to theses written and defended during 2010--2015 with
      Pearson's goodness-of-fit measure
      $(\text{E}-\text{O})^2/\text{E}$ between the expected and the
      observed numbers of marks awarded to theses written using
      \gls{tex}}
    \begin{tabularx}{\textwidth}{Xrrrrr}
      &\textbf{Without \TeX}&E(\textbf{With \TeX})
      &O(\textbf{With \TeX})&$(\text{E}-\text{O})^2/\text{E}$
      \\ \toprule
      \textbf{\parbox[t]{1em}{\centering A}} 
        &$15\,476$&$987.635$&\textcolor{OliveGreen}{$1\,181$}&
        $37.858$\\
      \textbf{\parbox[t]{1em}{\centering B}}
        &$9999$&$638.108$&\textcolor{red}{$587$}&$4.093$\\
      \textbf{\parbox[t]{1em}{\centering C}}
        &$7\,926$&$505.815$&\textcolor{red}{$381$}&$30.799$\\
      \textbf{\parbox[t]{1em}{\centering D}}
        &$4\,020$&$256.545$&\textcolor{red}{$194$}&$15.248$\\
      \textbf{\parbox[t]{1em}{\centering E}}
        &$2\,783$&$177.603$&\textcolor{red}{$128$}&$13.853$\\
      \textbf{\parbox[t]{1em}{\centering F}}
        &$1\,979$&$126.294$&\textcolor{OliveGreen}{$145$}&
        $2.771$\\
      \bottomrule
      \textbf{Total} &$\mathbf{42\,183}$&$\mathbf{2\,692}$&
        $\mathbf{2\,692}$&$\mathbf{104.623}$
    \end{tabularx}
    \label{table:statistics-contingency}
  \end{table}

  \begin{figure}
    \centering
    \begin{tikzpicture}
      \begin{axis}
        [
          ytick={1,2,3,4,5,6},
          xticklabels={,,A,B,C,D,E,F},
          yticklabels={With \TeX{} (at \acrshort{sci}),
                       Without \TeX{} (at \acrshort{sci}),
                       With \TeX{},
                       Without \TeX{},
                       With \TeX{} (at \acrshort{fi}),
                       Without \TeX{} (at \acrshort{fi})},
        ]
        \addplot+[
          boxplot prepared={
            lower quartile=1,
            median=1,
            upper quartile=2,
            upper whisker=3,
            lower whisker=1,
            average=1.9148387097
          },
        ] table[row sep=\\,y index=0] {
          data\\ 4\\ 5\\ 6\\
        };
        \addplot+[
          boxplot prepared={
            lower quartile=1,
            median=2,
            upper quartile=3,
            upper whisker=6,
            lower whisker=1,
            average=2.0201768307
          },
        ] table[row sep=\\,y index=0] {
          data\\
        };
        \addplot+[
          boxplot prepared={
            lower whisker=1,
            lower quartile=1,
            median=2,
            upper quartile=3,
            upper whisker=6,
            average=2.2110091743
          },
        ] table[row sep=\\,y index=0] {
          data\\
        };
        \addplot+[
          boxplot prepared={
            lower whisker=1,
            lower quartile=1,
            median=2,
            upper quartile=3,
            upper whisker=6,
            average=2.3971979233
          },
        ] table[row sep=\\,y index=0] {
          data\\
        };
        \addplot+[
          boxplot prepared={
            lower whisker=1,
            lower quartile=1,
            median=2,
            upper quartile=3,
            upper whisker=6,
            average=2.3565375303
          },
        ] table[row sep=\\,y index=0] {
          data\\
        };
        \addplot+[
          boxplot prepared={
            lower whisker=1,
            lower quartile=2,
            median=3,
            upper quartile=5,
            upper whisker=6,
            average=3.1209262436
          },
        ] table[row sep=\\,y index=0] {
          data\\
        };
      \end{axis}
    \end{tikzpicture}
    \caption{A box plot of the grades of theses written and
    defended during 2010--2015 at \acrshort{fi}, \acrshort{sci} and
    all the faculties of \gls{mu} with and without \gls{tex}}
    \label{fig:statistics-grades}
  \end{figure}

  Suppose the null \gls{hypothesis} $h_1$ that the grades awarded
  to theses written using and not using \gls{tex}, respectively,
  have the same distribution on the significance level
  $\alpha=0.05$. The one-tailed Pearson's $\chi^2$ test
  \cite{Pearson00} of the goodness of fit was applied to the
  observations of awarded grades that made up more than one percent
  of the sample data (grades A, B, C, D, E and F, see Table
  \ref{table:statistics-contingency}).
  Since \begin{equation}
    \sum_{\text{A},\text{B},\ldots,\text{F}}
    (\text{E}-\text{O})^2/\text{E}=104.623\gg11.07=
    \chi_{1-\alpha}^2(5)
  \end{equation} the author refused the null hypothesis $h_1$
  and concluded that the grades are indeed differently
  distributed on the significance level $\alpha$.

  Suppose also the null \gls{hypothesis} $h_2$ that the
  distribution of grade F being awarded to theses written using
  \gls{tex} is equivalent to the distribution of grade F being
  awarded to theses not written using \gls{tex}. The two-tailed
  Mann-Whitney U test \cite{mann47} was applied to the observations
  of grade F being and not being awarded to theses written using
  and not using \gls{tex}: \begin{align}
    \tag{Without \gls{tex}} n_1 &= 42\,183 \\
    \tag{With \gls{tex}}     n_2 &= 2\,692  \\
    U_1 &= 1\,979\cdot\left(145\cdot0.5+(2\,692-145)\cdot1\right) \\
    \nonumber&+(42\,183-1979)\cdot\left(145\cdot0+(2\,692-145)
      \cdot1\right) \\
      \nonumber&=56\,383\,784.5 \\
    U_2 &= 145\cdot\left(1\,979\cdot0.5+(42\,183-1\,979)\cdot1\right) \\
    \nonumber&+(2\,892-145)\cdot\left(1\,975\cdot0+(42\,183-1\,979)
      \cdot0.5\right) \\
      \nonumber&=57\,172\,851.5 \\
    U &= \min(U_1,U_2) = U_1 = 56\,383\,784.5
  \end{align} Since $n_1n_2\gg20$, $U\sim\text{N}(\frac{n_1n_2}2,
  \frac{n_1n_2(n_1n_2+1)}{12})$. After normalization to
  \begin{equation}
    \text{N}(0,1)\sim z =
    \frac{U-\frac{n_1n_2}2}{\sqrt{\frac{n_1n_2(n_1+n_2+1)}{12}}}
    \approx-\frac{394\,533.5}{651\,662.46}\approx-0.605
  \end{equation} the two-tailed \gls{pval} $\beta$ was computed by
  the author as follows:\begin{equation}
    \operatorname{arg\,min}\limits_{\beta} P(\Phi^{-1}_{\beta/2}
    \leq z\leq\Phi^{-1}_{1-\beta/2}) \approx 0.54518
  \end{equation}Since $\beta>\alpha$, the author concluded on the
  significance level $\alpha$ that there wasn't a strong evidence
  against the null \gls{hypothesis} $h_2$.
  
  On grounds of the refusal of $h_1$ and the inability to refuse
  $h_2$ the author concluded that the grades awarded to
  theses defended during 2010--2015 and written using \gls{tex}
  had been statistically significantly better than grades awarded
  to theses defended during 2010--2015 and not written using
  \gls{tex}. A box plot of the grades is shown in Figure
  \ref{fig:statistics-grades}.

  \section{Comparison with Existing Solutions}
  As a part of the initial research, I also assessed several
  templates for the typesetting of theses in order to gather ideas
  for immediate and future improvements of the \textsf{fithesis2}
  class.  Initially, I'm going to focus on templates used at Czech
  technical universities. I will then broaden the scope to include
  both the templates of foreign universities and other generic
  thesis typesetting \glspl{docclass}.

  \subsection{CTUStyle\index{CTUStyle|(} and
    CUStyle\index{CUStyle|(}}\label{sec:ctu&custyle}
  The first of the reviewed thesis typesetting templates were
  CTUStyle \cite{ctustyle} and CUStyle \cite{custyle}
  \glspl{texpackage} from RNDr. Petr Olšák, which were designed to
  be used at \gls{fel} and at \gls{cuni}, respectively. The sole
  dependencies of these templates are the \inx{\gls{csplain}} and
  \inx{\gls{opmac}} \glspl{texpackage}.

  The templates are closely tied with the visual styles of the
  universities, which is mainly achieved through color-coding.
  Aside from black-and-white text, the CTUstyle \gls{texpackage}
  typesets various typographic elements in the
  \clrpicker{pantone300} Pantone 300 color, which makes for a
  visually pleasing combination. The CUstyle \gls{texpackage} uses
  the combination of black, gray and \clrpicker{cured} Pantone 1797
  color to a similar end. By typesetting various typographic
  elements in the \clrpicker{pantone122} Pantone 122 color, which
  is a part of the visual identity of \gls{fi} \cite{filogo}, and
  by replacing the logotype of the faculty with its colored
  version, which was created as a part of Matúš Kominka's bachelor
  thesis \cite{Kominka08}, the \textsf{fithesis2} \gls{docclass}
  could be likewise revitalized.\partimp{v0.3.00, 01,
  06}\textsuperscript{,}\pending{Color \cs{tt} and section names.
  Define and document \cs{facultyheadings}, \cs{facultyhyperlinks},
  \cs{facultytables-light} and \cs{facultytables-dark} macros
  expanding to the names of colors of the given typographic
  elements.} Since each faculty of \gls{mu} has a color version of
  its logo \cite{muvis}, this redesign could be implemented without
  the loss of support for other faculties.\implemented{The logos were
  downloaded as monochromatic and direct color \gls{eps} files from
  \cite{muvis} and then transformed into \gls{pdf} files using the
  \texttt{epstopdf} \cite{epstopdf} utility.}

  Unlike with \textsf{fithesis2}, documents typeset with the
  CTUStyle and CUStyle templates are double-sided by default. The
  \gls{textwidth} of CTUStyle \eqref{eq:ctustyle-textwidth} and
  CUStyle \eqref{eq:custyle-textwidth}\begin{align}
    \label{eq:ctustyle-textwidth} 210\text{\,mm}\text{ (A4 width) }
      &- 2\cdot32\text{\,mm} \text{ \cite[line~249]{ctustyleCode} }
      &= 146\text{\,mm} \\
    \label{eq:custyle-textwidth}  210\text{\,mm}\text{ (A4 width) }
      &- 2\cdot31\text{\,mm} \text{ \cite[line~229]{custyleCode} }
      &= 148\text{\,mm}
  \end{align} is also much larger than that of \textsf{fithesis2}
  (127\,mm \cite[lines~989, 1017, 1045]{fithesis2Code}). The
  margins of CTUStyle and CUStyle, in conjunction with the chosen
  font family, \label{overlong-lines} allow for up to 100
  characters per line, which is not only wearying to the eye of an
  inexperienced reader \cite[section~2.1.2]{eletypostyle}, but also
  against the recommendations of \gls{fi}
  \cite[section~3.2.3]{bpdpfi}. As stated by the author of the
  styles \cite[post~25]{ctustyle-forums}, both decisions have
  likely been made in order to reduce the required storage space
  for the archival of physical prints of theses.\index{CTUStyle|)}
  \index{CUStyle|)}

  \subsection{Felthesis\index{felthesis}}\label{sec:felthesis}
  Being a traditional \gls{docclass}, felthesis from Ing. Vít Zýka
  Ph. D. provides an alternative for those \gls{fel} students, who
  prefer \gls{latex} over \gls{csplain} used by the CTUStyle
  \gls{texpackage}. Much like \textsf{fithesis1} and
  \textsf{fithesis2}, felthesis loads the KOMA-Script scrreprt
  \gls{docclass}, redefines some of the commands and \glspl{env},
  loads additional packages and defines additional thesis
  sectioning commands. Unlike \textsf{fithesis1} and
  \textsf{fithesis2}, felthesis loads babel and then selects the
  language based on the chosen language of the thesis
  \cite[lines~687--691]{felthesisCode}\pending{Load babel or
  polyglossia by default and pick the language based on
  \cs{@thesislang}}. Felthesis also loads Bib\LaTeX\pending{}
  \cite[line~722]{felthesisCode}, automatically generates
  index\pending{} \cite[line~763]{felthesisCode} and stamps the
  title, author, subject, and keywords into the header of the
  resulting \gls{pdf} file \cite[lines~959--971]{felthesisCode}
  \label{pdf-stamping}\partimp{v0.3.02}\textsuperscript{,}
  \pending{Turn the \cs{ThesisKeyWords} \gls{env} into a command.
  Add a command or additional argument to provide keywords in
  English. These will be both typeset and stamped into the header
  of the resulting \gls{pdf} file.}. Alongside the package
  \gls{ltxpackage}, both a technical \cite{felthesisCode} and a
  user documentation \cite{felthesis} is
  distributed\pending{}.

  \subsection{The Charles University in Prague}
  Along with Petr Olšák's CUStyle, \gls{cuni} also provides an
  official \gls{latex} template for the typesetting of theses
  \cite{cunisablona}. Rather than defining a new \gls{docclass},
  the archive contains a \gls{makefile} and a skeleton \gls{latex}
  document using the base report \gls{docclass} for both Czech and
  English theses. The students then modify said document to suit
  their requirements. The template uses \gls{cslatex} rather than
  the babel \gls{ltxpackage}, which means that \gls{csfonts} rather
  than the superior \gls{lmfonts} \cite{cslatexvsbabel} are used by
  default.

  \subsection{Mendel University in Brno}\index{\texttt{dipp.sty}|(}
  \label{sec:dipp.sty}
  Despite not being an official thesis template of \gls{mendelu},
  \texttt{dipp.sty} from doc. Ing. Dr. Jiří Rybička
  \gls{ltxpackage} warrants a mention. The package depends on the
  extended macro set of the \Hologo{XeTeX}
  \glslink{texeng}{typesetting engine} and is intended to be used
  with the base article \gls{docclass}. Although the article class
  was not designed with European typography in mind, the
  \texttt{dipp.sty} package redefines much of the geometry.

  Noteworthy is also the rich selection of additional markup, which
  is meant to ease the task of typesetting a thesis for those
  unfamiliar with \gls{latex}. To this end, thesis sectioning
  commands, macros and wrapper \glspl{env} for the inclusion of
  bibliography, tables or figures and discretionary macros specific
  to Czech typography such as \cs{az} for the typesetting of ranges
  and \cs{spoj} for hyphenation
  \cite{dippman}.\index{\texttt{dipp.sty}|)}

  \subsection{Brno University of Technology}
  \index{\texttt{thesis.sty}|(}
  The official thesis typesetting template of \gls{vut} is the
  \gls{ltxpackage} \texttt{thesis.sty}. Unlike
  \texttt{\hyperref[sec:dipp.sty]{\tt dipp.sty}},
  \texttt{thesis.sty} is meant to be used with the base report
  \gls{docclass}. All packages required by the \gls{ltxpackage} are
  documented within \cite[p.~9]{thesis.sty-doc}\pending{}. The
  implicit \gls{textwidth} of the \gls{ltxpackage} is 150\,mm
  \cite[p.~2]{thesis.sty-doc}, which, in conjunction with the usage
  of \gls{csfonts} at 12\,pt, allows for up to 90 characters per
  line. This is problematic for reasons mentioned in Section
  \ref{overlong-lines}.

  The \texttt{thesis.sty} \gls{ltxpackage} also adds provides
  additional markup. This includes wrapper commands such as
  \cs{novazkratka}, \cs{zk}, \cs{zkratka} and \cs{zkratkatext} for
  the typesetting of acronyms, the \cs{seznamzkratek} and
  \cs{literatura} \glspl{env} for acronyms and bibliography,
  respectively, and various math mode macros (see
  \cite[p.~6--9]{thesis.sty-doc}). Like
  \hyperref[sec:felthesis]{felthesis}, \texttt{thesis.sty} is also
  able to stamp thesis metadata into the \gls{pdf} header. The
  \texttt{thesis.sty} \gls{ltxpackage} can be downloaded from
  \cite{thesis.sty}.
  \index{\texttt{thesis.sty}|)}

  \subsection{Technical University of Liberec}
  \index{tul|(}\index{tulthesis|(}
  Unlike the preceding typesetting templates, tulthesis from doc.
  RNDr. Pavel Satrapa, Ph.D. is not a standalone \gls{docclass},
  but rather a part of the tul \gls{ltxpackage}, which is the
  official \gls{ltxpackage} used for the typesetting of documents
  on \gls{tul}. The tulthesis \gls{docclass} uses the base report
  \gls{docclass} \cite[section~1]{tulthesis-man} and it is very
  visually distinctive. It inspired \cite[post~21]{ctustyle-forums}
  the creation of Petr Olšák's \nameref{sec:ctu&custyle}
  \glspl{texpackage}. The tul \gls{ltxpackage}, along with the
  tulthesis \gls{docclass}, can be downloaded from \cite{tul}.
  \index{tul|)}\index{tulthesis|)}

  \subsection{Technical University of Ostrava}\index{diploma|(}
  The official style of \gls{vsb} is the diploma \gls{docclass} by
  Jiří Dvorský. Like Jiří Rybička's \hyperref[sec:dipp.sty]{\tt
  dipp.sty}, diploma uses the base article \gls{docclass}, modifies
  its geometry and adds thesis sectioning commands and additional
  macros for the users' convenience \cite{diploma}. The output
  documents are one-sided by default and the font size of 11\,pt is
  too small for the \gls{textwidth} of \begin{equation}
    \begin{split}
      210\,\text{mm} \text{ (A4 width) }
        &- 28\,\text{mm} \text{ (left margin) }  \\
        &- 32\,\text{mm} \text{ (right margin) } = 150\,\text{mm}
    \end{split}
  \end{equation} as defined in
  \cite[lines~111,~123]{diplomaSource}, resulting in overlong lines
  as discussed earlier (see section \ref{overlong-lines}).

  The additional macros range from mathematical \glspl{env}
  \cite[section~3.5]{diploma} and commands and \glspl{env} for the
  typesetting of code with syntax highlighting
  \cite[section~3.6]{diploma}\pending{} through commands for the
  insertion of graphics \cite[section~3.7]{diploma}.
  \index{diploma|)}

  \subsection{Silesian University in Opava}
  At the time of writing this thesis, there is no official or
  unofficial \gls{gls-docclass} or \gls{ltxpackage} of \gls{slu}
  available on the internet. There is a mention \cite{vavreckova}
  of the works on a \gls{docclass} at the webpage of RNDr. Šárka
  Vavřečková, Ph.D. who teaches typography at \acrlong{fpf}. There
  is also a bachelor's thesis \cite{hanzal09} written by Michal
  Hanzal from \gls{opf}. The author of the thesis describes the
  process of creating a \gls{gls-docclass} for what is supposed to
  be the official thesis typesetting template at \gls{opf}, but
  there is no mention of the style at the website of the faculty
  \cite{opftemplates}, although the linked dean's directives were
  written with \gls{latex} in mind.

  \subsection{Palacký Univerzity Olomouc}
  There are three independant \gls{latex} thesis typesetting
  templates available at \acrlong{upolsci} for the students at the
  departments of Experimental Physics, Informatics and Mathematical
  Analysis and Applications, respectively.

  The template used at the Department of Experimental Physics is
  neither a \gls{docclass} nor a \gls{ltxpackage}. All definitions
  and alterations to the used built-it article \gls{docclass} are
  stored directly inside the example \gls{latex} source code, which
  the students are supposed to edit according to the needs of their
  theses. The template is supposed to be typeset using
  \glslink{cslatex}{\pdfCSLaTeX{}} \gls{texeng} meaning that
  \gls{csfonts} instead of the superior \gls{lmfonts} are used.
  Along with the thesis template, a \gls{beamer} template for
  thesis defense presentations is also distributed\pending{}. Both
  templates can be downloaded at \cite{upol-exfyz-template}.

  \index{KIdiplom|(} The template used at the Department of
  Informatics is KIdiplom from Jan Outrata and Martin Rotter and it
  is a combination of a \gls{docclass}, which is based on the base
  article \gls{docclass} and is meant for the typesetting of
  theses, and a \gls{ltxpackage}, which was designed as a general
  \gls{ltxpackage} for the typesetting of documents in the visual
  style of the Department of Informatics. Amongst the vices of the
  template are the disabled double-sided typing
  \cite[lines~69--72]{upol-inf-templateSource} and the
  \gls{textwidth} of\begin{equation}
    426\text{\,pt}\text{ (\texttt{a4paper} \gls{textwidth}) }
    \cdot1{,}05\text{ \cite[line~81]{upol-inf-templateSource} }
    \approx157,23\text{\,mm}
  \end{equation} in conjunction with the 12\,pt \gls{lmfonts} is
  too generous \cite[lines~74--82]{upol-inf-templateSource} causing
  overlong lines (see section \ref{overlong-lines}). The template
  by default loads \gls{biblatex} and \glspl{ltxpackage} for the
  typesetting of glossaries and index. The added convenience macros
  then comprise thesis sectioning commands and \glspl{env} for the
  typesetting of mathematical definitions, theorems or proofs and
  code snippets. Along with the thesis template, \gls{latex}
  templates for the typesetting of thesis reader's
  reports\pending{See the templates by \MR} and \gls{biblatex}
  ISO\hyph{}690 \cite{iso690-1,iso690-2} citation styles\pending{}
  are distributed. The templates can be downloaded at
  \cite{upol-inf-template}.\index{KIdiplom|)}

  \index{\texttt{maam-dipl.sty}|(}The template used at the
  Department of Mathematical Analysis and Applications is a simple
  \gls{ltxpackage} \texttt{maam-dipl.sty}, which only defines basic
  thesis sectioning commands and is meant to be used with the base
  article \gls{docclass}. The template can be downloaded at
  \cite{upol-maam-template}\index{\texttt{maam-dipl.sty}|)}.

  % VUTBr

  \subsection{Foreign Universities}
  \subsection{Generic templates}

  % CTAN: http://www.ctan.org/topic/dissertation
  % Kumar07, https://tug.org/pracjourn/2007-4/

  \chapter{Implementation}
    \section{Design goals}
    \section{Architecture}
    \section{\Hologo{XeLaTeX} Support}
  % \section{OpenType Font Support}
  \chapter{Closing Notes}

  % Bibliography
  \newpage
  \nocite{*}
  \tolerance=300
  \emergencystretch=1em
  \printbibliography[heading=bibintoc]
  \par

  % Glossary
  \newpage
  \printglossaries

  % Index
  \printindex

  % Appendices
% \begin{appendices}
%  \chapter{Example Theses}
%  \chapter{Technical Documentation}
%  \includepdf[pages=-]{fithesis3/fithesis3/fithesis.pdf}
%  \chapter{User Documentation}
%  \includepdf[pages=-]{fithesis3/fithesis3/guide/mu/fi.pdf}
%  \chapter{Thesis writing guidelines}
% \end{appendices}

\end{document}
