\documentclass[12pt]{beamer}
% Packages
\usepackage{polyglossia}
\setmainlanguage{english}
% Theme
%% Fonts
\usepackage{fontspec}
\setsansfont[Ligatures=TeX]{Comenia Sans}
%% Colors
\usetheme{Boadilla}
%% Miscellanea
\frenchspacing
\beamertemplatenavigationsymbolsempty
%% Nested itemize fix
%% http://tex.stackexchange.com/a/74069/70941
\usepackage{etoolbox}
\setbeamercovered{transparent}
\makeatletter
\newcommand*\fix@beamer@close{%
  \ifnum\beamer@trivlistdepth>0%
    \beamer@closeitem%
  \fi}
\newcommand*\fix@beamer@open{%
  \ifnum\beamer@trivlistdepth>0%
    \gdef\beamer@closeitem{}%
  \fi}
\BeforeBeginEnvironment{enumerate}{\fix@beamer@close}
\AfterEndEnvironment{enumerate}{\fix@beamer@open}
\BeforeBeginEnvironment{itemize}{\fix@beamer@close}
\AfterEndEnvironment{itemize}{\fix@beamer@open}
\BeforeBeginEnvironment{description}{\fix@beamer@close}
\AfterEndEnvironment{description}{\fix@beamer@open}
\makeatother
% Metadata
\title{The Form of Theses\\ Written in \LaTeX}
\subtitle{Thesis defense}
\date{May 28, 2015}
\author{Vít Novotný}
\institute[FI MU]{Faculty of Informatics,\\Masaryk University in Brno}
\begin{document}
  \frame{\maketitle}
  \begin{frame}
    \frametitle{History}
    \begin{itemize}[<+->]
      \item The \emph{fithesis1} and \emph{fithesis2} document classes
      \item The \emph{xslt2} module
      \item The \emph{comenia} package
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{\TeX\ usage at MUNI}
    \begin{itemize}[<+->]
      \item User survey
      \begin{itemize}
        \item Most respondents were FI bachelor's degree students.
        \item Most preferred \LaTeX\ over the alternatives.
        \item Some submitted feedback regarding \emph{fithesis1} and \emph{2}.
      \end{itemize}
      \item Analysis of the 45\,488 theses defended during 2010--2015:
      \begin{itemize}
        \item A heuristic was devised to detect the usage of \TeX.
        \item \TeX\ usage:
          \begin{itemize}
            \item Was dominated by FI (59.09\%) and Sci (14.9\%).
            \item Correlated with better grades.
            \item Steady 0.6\% increase overall across faculties
            \item Steady 10\% decrease overall in rig. and doc. theses at Sci
          \end{itemize}
      \end{itemize}
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{Review of Existing Templates}
    \begin{itemize}[<+->]
      \item Moderate use of color (\emph{CTUStyle, CUStyle, tul})
      \item Separation of presentation and logic (\emph{tul+tulthesis})
      \item Additional mark-up (\texttt{dipp.sty}, \texttt{thesis.sty}, \emph{diploma})
      \item Homebrew \textsc{Bib}\LaTeX\ ISO-690 citation styles (\emph{KIdiplom})
      \item Common problems:
      \begin{itemize}
        \item Lack of double-sided typesetting
        \item Overly large \texttt{\string\textwidth}
      \end{itemize}
    \end{itemize}
    \uncover<8>{Some features had a direct impact on the development of \emph{fithesis3}, the rest is a good starting point for the future development of the class.}
  \end{frame}
  \begin{frame}
    \frametitle{Making the class modular: Rationale, Design and Implementation}
    \begin{itemize}[<+->]
      \item \emph{Fithesis1} and \emph{2} mix logic, locales and presentation.
      \item Addition of new locales and styles is non-trivial.
      \item \emph{Fithesis3} is composed of the following parts:\begin{enumerate}
        \item\emph{Locale files} that encapsulate locale strings.
          \begin{itemize}
            \item Czech, English and Slovak are supported by default.
            \item New locales can be easily added.
          \end{itemize}
        \item\emph{Style files} that define the look and layout for each faculty.
          \begin{itemize}
            \item Prepared in accordance with the Dean's Directives of the respective faculties.
          \end{itemize}
        \item\emph{The document class} that processes the user input and loads
          the according locale and style files.
      \end{enumerate}
      \item Each of these parts is extensively documented.
      \item Much of the feedback from the user survey and the features of existing templates were implemented.
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{Other changes}
    \begin{itemize}[<+->]
      \item Metafont $\rightarrow$ vector graphics transition
      \item Better metadata support
      \item \ldots
    \end{itemize}
  \end{frame}
  \begin{frame}
    \frametitle{Conclusion}
    The first paragraph of the respective chapter.
  \end{frame}
  \begin{frame}
    \frametitle{Future work}
    \begin{itemize}[<+->]
      \item Testing will continue until the end of June.
      \item The class will then be released at:\begin{itemize} 
        \item Online collaboration platforms (Share\LaTeX, Overleaf)
        \item The CTAN archive (for the inclusion in \TeX\ distibutions)
      \end{itemize}
      \item The class has been added to the curriculum of the \emph{FI:PB029 -- Electronic Document Preparation} subject
      \item During July--September 2015:\begin{itemize} 
        \item User documentation will be expanded.
        \item An author’s cookbook will be created.
        \item Thesis defense and thesis report templates will be released.
      \end{itemize}
    \end{itemize}
  \end{frame}
\end{document}
